Superior paper quality with R-Markdown: The OBVIOUS Study

### 2024-05-12 doc/manuscript.Rmd
more lightweight analysis

Your name here, your degree,^1^ 
Second in Command, MS,^1^ 
Minor Contributor, PhD,^2^ and 
The Mastermind, PhD^1^

^1^Your department, University, City, State. `r single_par`
^2^Department of Win, Overtly Pompous University, Irvine, CA. `r single_par`

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

**Correspondence to:** \
Your name, Degree \
Your Department \
Additional departmental details if needed \
Your university \
Address line 1 \
Address line 2 \
Address line 3 \
your phone number \
your email `r single_par`

**Word Count:** Abstract: ; Manuscript:

```{r data-prep, include=FALSE}

library(palmerpenguins)

data <- penguins

# ABSTRACT

# INTRODUCTION

Here's how to cite one thing.[@r_language]

Here's how to cite two or more things.[@ambrosius_design_2014; @sprint_research_group_randomized_2015]

# METHODS

# RESULTS

```{r results-prep}

n_penguins <- nrow(data)

mean_bill_length <- mean(data$bill_length_mm, na.rm = TRUE)

mean_mass <- mean(data$body_mass_g, na.rm = TRUE)
sd_mass <- sd(data$body_mass_g, na.rm = TRUE)

# make a short name for table.glue::table_value - this makes
# the inline reporting of results less complicated to read.
tv <- table.glue::table_value
tg <- table.glue::table_glue

Here's how to put some results into the text: A total of `r n_penguins` were included in the current analysis. The mean bill length was `r tv(mean_bill_length)`. The mean (standard deviation) body mass was `r tg("{mean_mass} ({sd_mass})")`

# DISCUSSION

# REFERENCES

<div id="refs"></div>

# ACKNOWLEDGMENTS

# FINANCIAL DISCLOSURE

```{r tbl_footers}

ckd_defn <- as_paragraph("Chronic Kidney Disease was defined by an estimated glomerular filtration rate <60 ml/min/1.73 m", as_sup(2L), " based on the 2021 CKD-EPI creatinine equation.")

@@

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  MoCA = 'Montreal cognitive assessment',
  EHR  = 'electronic health records',
  CVD  = 'cardiovascular disease'
)

@@

**Table 1: Penguin characteristics**

tbl_summary(data, by = 'island') %>% 
 as_flex_table() %>% 
 theme_box() %>% 
 align(j = c(2,3,4), align = 'center')

@@

**Table 2: A more customized table**

@@

data %>% 
 group_by(species) %>% 
 slice(1:3) %>% 
 select(-body_mass_g, -sex, -year) %>% 
 mutate(across(where(is.numeric), table_value)) %>% 
 as_grouped_data(groups = 'species') %>% 
 as_flextable(hide_grouplabel = TRUE) %>% 
 set_header_labels(bill_length_mm = 'Bill length, mm',
                   bill_depth_mm = 'Bill depth, mm',
                   flipper_length_mm = 'Flipper length, mm') %>% 
 add_header_row(values = c(" ", " Penguin things"),
                colwidths = c(1, 3)) %>% 
 bg(i = ~!is.na(species), bg = 'grey90') %>% 
 width(width = 1.5) %>% 
 width(j = 1, width = 2) %>% 
 align(part = 'all', j = c(2,3,4), align = 'center') %>% 
 footnote(i = 2, j = 2, 
          part = 'header',
          value = as_paragraph("Bill length doesn't correlate with what you think it does."),
          ref_symbols = '1') %>% 
 footnote(i = 1, j = 1, 
          part = 'body',
          value = as_paragraph('Adelies are the best and they know it'), 
          ref_symbols = '2') %>% 
 add_footer_lines(values = abbrvs_write(abbrvs[c('CI', 'EHR', 'SD')])) %>% 
 add_footer_lines(values = "none of those abbreviations are in the table but they demonstrate how to include abbreviations using my helper functions")

@@

<!-- new page not needed at the end of a block section -->

@@

margins_normal <- page_mar(bottom = 2.54 * 0.3937, 
                           top    = 2.54 * 0.3937, 
                           right  = 2.54 * 0.3937, 
                           left   = 2.54 * 0.3937,
                           gutter = 0)

block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous",
    page_margins = margins_normal
  )
)

@@

To make a landscape section, use the block_section code here. Note that the section will be applied to all content prior to the block reaching up to the most recent block.

```{r echo = TRUE}
block_section(
  prop_section(
    page_size = page_size(orient = "landscape"),
    page_margins = page_mar(bottom = 1/8, top = 1/8, 
                            right = 1/4, left = 1/4),
    type = "continuous"
  )
)
```

To switch back to regular portrait section, add another portrait block. Notice how the block below acts on all content above it up to the prior block.

```{r echo = TRUE}
block_section(
  prop_section(
    page_size = page_size(orient = "portrait"),
    type = "continuous",
    page_margins = margins_normal
  )
)
```

>>>>

withr::with_dir(
  here::here(),
  code = {
    targets::tar_load(c(labels,
                        analysis,
                        tbl_exclusion,
                        tbl_characteristics,
                        tbl_cuminc))
  }
)

@@

abbrvs <- list(
  CI   = 'confidence interval',
  FI   = 'frailty index',
  SD   = 'standard deviation',
  IQR  = 'interquartile range',
  CVD  = 'cardiovascular disease',
  DPP  = "Diabetes Prevention Program",
  MESA = "Multi-Ethnic Study of Atherosclerosis"
)

@@

fn_mesa_baseline <- as_paragraph("Baseline for Multi-Ethnic Study of Atherosclerosis participants was visit 1 or 2 dependent on pre-diabetes status")

ft_tidy <- function(x){
  theme_box(x) %>% 
    align(align = 'center', part = 'all') %>%
    align(j = 1, align = 'left', part = 'all')
}

@@

Table 0: Participants included.

```{r tbl_exclusion}

tbl_exclusion %>% 
  flextable() %>% 
  width(j = 'name', width = 3) %>% 
  add_footer_lines(
    values = abbrvs_write(abbr = abbrvs[c("DPP", "MESA")])
  ) %>% 
  set_header_labels(name = "Inclusion criteria") %>% 
  footnote(part = 'body', 
           i = 2, 
           j = 'name', 
           ref_symbols = '1', 
           value = fn_mesa_baseline) %>% 
  ft_tidy()

Table 1: Descriptive table of Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis participants included in the current analysis

@@

tbl_characteristics$included %>% 
  as_flex_table() %>% 
  ft_tidy()

@@

Table 2: Descriptive table of Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis participants excluded from the current analysis

@@

tbl_characteristics$excluded %>% 
  as_flex_table() %>% 
  ft_tidy()

@@

Table 3. Cumulative incidence of diabetes over follow-up in the Diabetes Prevention Program and Multi-Ethnic Study of Atherosclerosis

tbl_cuminc %>% 
  ft_tidy() %>% 
  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')

Table 4.

analysis$eval %>% 
  pivot_wider(names_from = evaluation, 
              values_from = c(standard, individualized)) %>% 
  select(stat, 
         standard_internal,
         individualized_internal,
         standard_external,
         individualized_external) %>% 
  separate(stat, into = c("group", "stat"), sep = '\\.\\.', fill = 'left') %>% 
  mutate(
    stat = recode(
      stat,
      demo_parity = "Demographic parity",
      equal_oppo = "Equal opportunity",
      equal_odds = "Equal odds"
    ),
    group = recode(
      group,
      race = "Race/ethnicity",
      sex = "Sex"
    )
  ) %>% 
  as_grouped_data(group = 'group') %>% 
  janitor::remove_empty('rows') %>% 
  as_flextable(hide_grouplabel = TRUE) %>%
  set_header_labels(
    stat = "Evaluation statistic",
    standard_internal = "Standard",
    individualized_internal = "Individualized",
    standard_external = "Standard",
    individualized_external = "Individualized"
  ) %>% 
  add_header_row(values = c("Evaluation statistic", 
                            "Internal Validation", 
                            "External Validation"), 
                 colwidths = c(1, 2, 2)) %>% 
  ft_tidy() %>% 
  width(width = 1.5) %>% 
  merge_v(part = 'header', j = 1) %>% 
  bg(i = ~ !is.na(group), part = 'body', bg = 'grey80')
